# Data Description
```{r setup, include=FALSE}
#| message: false
 
source("load_data.R")
library(glue)

descriptions <- list(
  "UNITID" = "Department of Education assigned identifier.",
  "CDR3" = "The 3-year cohort default rate.  The 3-year cohort default rate is the percentage of a school's borrowers who enter repayment on certain Federal Family Education Loan (FFEL) Program or William D. Ford Federal Direct Loan (Direct Loan) Program loans during a particular federal fiscal year (FY), October 1 to September 30, and default or meet other specified conditions prior to the end of the second following fiscal year. Reported as a proportion ([0-1]).",
  "INSTNM" = "The name of the institution.",
  "HBCU" = "Self reported status as Historicaly Black College or University (HBCU). Reported as a 0/1 indicator",
  "MENONLY" = "Institution only enrolls men. Reported as 0/1 indicator",
  "WOMENONLY" = "Institution only enrolls women. Reported as 0/1 indicator",
  "RELAFFIL" = "Reported religious affiliation of the insitution. Reported as a categorical value:
    - -1 (Not Reported)
    - -2 (Not Applicable)
    - Positive integer (indicates type of affilation, details in data dictionary).",
  "DISTANCEONLY" = "Insitution exlusively provides distance-only education. Reported as 0/1 indicator.",
  "CONTROL" = "Control of Institution. Reported as a categorical value:
    - 1 Public
    - 2 Private, Non-Profit
    - 3 Private, For-Profit
  
    We have elected to summarize this as Public vs. Private given the small number of For-Profit schools in our dataset.
  ",
  "LOCALE" = "Locate of Instutution. Reported as a categorical value:
  
  - 11	City: Large (population of 250,000 or more)
  - 12	City: Midsize (population of at least 100,000 but less than 250,000)
  - 13	City: Small (population less than 100,000)
  - 21	Suburb: Large (outside principal city, in urbanized area with population of 250,000 or more)
  - 22	Suburb: Midsize (outside principal city, in urbanized area with population of at least 100,000 but less than 250,000)
  - 23	Suburb: Small (outside principal city, in urbanized area with population less than 100,000)
  - 31	Town: Fringe (in urban cluster up to 10 miles from an urbanized area)
  - 32	Town: Distant (in urban cluster more than 10 miles and up to 35 miles from an urbanized area)
  - 33	Town: Remote (in urban cluster more than 35 miles from an urbanized area)
  - 41	Rural: Fringe (rural territory up to 5 miles from an urbanized area or up to 2.5 miles from an urban cluster)
  - 42	Rural: Distant (rural territory more than 5 miles but up to 25 miles from an urbanized area or more than 2.5 and up to 10 miles from an urban cluster)
  - 43	Rural: Remote (rural territory more than 25 miles from an urbanized area and more than 10 miles from an urban cluster)
  
  We have elected analyze this variable as City, Suburb, Town, and Rural.",
  "ADM_RATE" =  "The admission rate. This is the number of undergraduate applicants who are offerred admission divided by the total number of undergraduates who applied in the Fall. Reported as a proportion ([0-1])",
  "SATVRMID" = "Median SAT Verbal Subsection Score for Admitted Students. Reported as an integer.",
  "SATMTMID" = "Median SAT Math Subsection Score for Admitted Students. Reported as an integer.",
  "SATWRMID" = "Median SAT Writing Subsection Score for Admitted Students. Reported as an integer.",
  "UGDS" =  "The number of undergraduate degree and certificate seeking students. Reported as an integer.",
  "TUITIONFEE_IN" = "Total tuition and required feeds for in-state students. Reported in dollars.",
  "TUITIONFEE_OUT" = "Total tuition and required feeds for out-of-state students. Reported in dollars.",
  "INEXPFTE" = "Instructional expenditures per full time equivalent student. Reported in dollars.",
  "AVGFACSAL" = "Average faculty salary per month. Computed by dividing the total salary outlays by the number of months worked for all full-time, nonmedical instructional staff. Reported in dollars.",
  "PFTFAC" = "Proportion of faculty that are employed full-time. Reported as a proportion([0-1]).",
  "PCTPELL" = "Proportion of undergraduates who recieved a Pell grant in that academic year. Reported as a proportion ([0-1]).",
  "PCTFLOAN" =  "Proportion of undergraduates who recieved a federal student loan in that academic year. Reported as a proportion ([0-1]).",
  "RET_FT4" = "Proportion of full-time, first-time, degree/certificate-seeking undergraduates who were enrolled at the institution in the fall 1 year after starting at the institution. Reported as a proportion ([0-1]).",
  "GRAD_DEBT_MDN" = "Median debt at graduation for students who complete their program. Reported in dollars.",
  "HIGH_CDR" = "(Derived) Indicator variable for schools with a > 10% 3 year-default rate(`CDR3 > 0.1`)",
  "SINGLEGENDER" = "(Derived) Indicator variable for schools with a single gender (`MENONLY = 1 OR WOMENONLY = 1`)",
  "RELIGIOUS" = "(Derived) Indicator variable for schools with a religious affiliation (`RELAFFIL > 0`)"
)
```

## Data Source
Our data comes from the [US Department of Education College Scorecard](https://collegescorecard.ed.gov/), specifically the ["Most Recent Institution-Level Data" from September 14, 2022](https://ed-public-download.app.cloud.gov/downloads/Most-Recent-Cohorts-Institution_09012022.zip). The raw dataset contains 2,989 features on 6,681 institutions but we've chosen to focus on 1,990 institutions that are:

-   Primarily four-year degree granting institutions (for undergraduates)
-   The main campus only (if an overall institution has multiple campuses)

[This dataset is well documented](https://collegescorecard.ed.gov/data/documentation/) and [we have built a small pipeline for ingesting the raw DOE data and creating a dataset for our analysis](https://github.com/ajb5d/CollegeScorecard/blob/main/get_data.R). 

## Variables



```{r}
#| output: asis
#| echo: false

for (var in names(data)) {
  description <- ""
  if (!is.null(descriptions[[var]])) {
    description <- descriptions[[var]]
  }
  print(glue("* `{var}`: {description}"))
}
```

