# Linear Regression

```{r setup, include=FALSE}
library(MASS)
library(tidyverse)
library(GGally)
library(patchwork)
source("load_data.R")
```

## Introduction
The question for linear regression is what factors influence in-state tuition at U.S. universities. We will examine the factors influencing the total in-state tuition and fees for main campus, four year, degree granting institutions.

Our primary response variable is `TUITIONFEE_IN` -- the total tuition and fees for in-state students in US dollars.

The cost of attending college is one of the [main concerns for high school students planning for undergraduate studies](https://studentloanhero.com/featured/undergrad-worries-survey/). Tuition and fees have significantly increased over the past decade which resulted in [increased student debts and an increase in student loans](https://educationdata.org/student-loan-debt-crisis):

-   Total \$1.75 trillion in student loan debt (including federal and private loans)
-   On average, each borrower owes \$28,950
-   About 92% of all student debt are federal student loans
-   More than half (55%) of students from public four-year institutions had student loans

With the above facts, it is important that we understand the factors that influence high tuition fees. These factors that impact the fees needs to be addressed to encourage the current and future generations can have a debt-free college education of their choice.

## Visualizations
We started off by generating scatter plots (@fig-corr) across the various variables to help show and provide insight to some initial relationships. None of the predictor variables display an obvious linear relation ship to in-state tuition during this first pass of visualization. There are some predictors where the scatter plot looks to be creating a "C" shape. We will explore those variables further while also layering in some of those categorical variables later.

```{r, echo=FALSE}
#| label: fig-corr
#| fig-height: 10
#| fig-width: 7.5
#| fig-cap: "Correlations of Continuous Variables"

custom_theme <- theme_bw() + theme(
  axis.text = element_text(size = 6),
  strip.text.x = element_text(size = 6),
  strip.text.y = element_text(size = 6))

cor1 <- ggpairs(data[,c("TUITIONFEE_IN","ADM_RATE","UGDS","INEXPFTE","AVGFACSAL","PFTFAC")], 
                lower = list(continuous = wrap("points", size = 0.1, alpha = 0.3)),
                progress = FALSE) + custom_theme
cor2 <- ggpairs(data[,c("TUITIONFEE_IN","PCTPELL","PCTFLOAN","RET_FT4","GRAD_DEBT_MDN")], 
                lower = list(continuous = wrap("points", size = 0.1, alpha = 0.3)),
                progress = FALSE) + custom_theme

wrap_elements(ggmatrix_gtable(cor1)) / wrap_elements(ggmatrix_gtable(cor2))
```


We illustrate the impact of `CONTROL` on the relationship between various predictor variables and in-state tuition(@fig-cost-by-cont) and this helps explain the "C" shape we were seeing in some of the earlier scatter plots, as the difference between Public and Private schools. It appears the estimated slopes for some of these relationships will vary depending on whether the school is public or private. Specifically it looks like there may be a negative relation ship between `TUITIONFEE_IN` and `ADM_RATE`, `PCTPELL`, and `PCTFLOAN` when `CONTROL` is private. It also looks like there may be a positive relationship between `TUITIONFEE_IN`and `AVGFACSAL` and `RET_FT4` that is greater when `CONTROL` is private rather than public.

```{r, echo=FALSE}
#| echo: false
#| fig-cap: "Tuition versus Continuous Variables, by Control"
#| label: fig-cost-by-cont
#| fig-height: 10
#| fig-width: 7.5
admscp <- data %>%
  select(ADM_RATE, TUITIONFEE_IN, CONTROL) %>% 
  drop_na() %>% 
  ggplot(aes(x = ADM_RATE, y = TUITIONFEE_IN, color = CONTROL)) +
    scale_x_continuous(labels = scales::label_percent()) + 
    scale_y_continuous(labels = scales::label_dollar()) + 
    geom_point(alpha = 0.5) +
    geom_smooth(method = "lm", se = FALSE) +
    labs(
      subtitle = "Tuition vs. Admission Rate",
      x = "Admission Rate",
      y = "In-State Tuition",
      color = "Insitution Type"
    ) + theme_bw()

ugdsscp <- data %>%
  select(UGDS, TUITIONFEE_IN, CONTROL) %>% 
  drop_na() %>% 
  ggplot(aes(x = UGDS, y = TUITIONFEE_IN, color = CONTROL)) +
    scale_x_continuous(labels = scales::label_comma()) + 
    scale_y_continuous(labels = scales::label_dollar()) + 
    geom_point(alpha = 0.5) +
    geom_smooth(method = "lm", se = FALSE) +
    labs(
      subtitle = "Tuition vs. Undergrad Students",
      x = "Undergrad Students",
      y = "In-State Tuition",
      color = "Insitution Type"
    ) + theme_bw()

avgfacsalscp <- data %>%
  select(AVGFACSAL, TUITIONFEE_IN, CONTROL) %>% 
  drop_na() %>%
  ggplot(., aes(x = AVGFACSAL, y = TUITIONFEE_IN,  color = CONTROL)) +
    scale_x_continuous(labels = scales::label_dollar()) +
    scale_y_continuous(labels = scales::label_dollar()) + 
    geom_point(alpha = 0.5) +
    geom_smooth(method = "lm", se = FALSE) +
    labs(
      subtitle = "Tuition vs. Average Faculty Salary",
      x = "Avg. Faculty Salary (USD per month)",
      y = "In-State Tuition",
      color = "Insitution Type"
    ) + theme_bw()

pftfacscp <- data %>%
  select(PFTFAC, TUITIONFEE_IN, CONTROL) %>% 
  drop_na() %>%
  ggplot(., aes(x = PFTFAC, y = TUITIONFEE_IN,  color = CONTROL)) +
    scale_x_continuous(labels = scales::label_percent()) +
    scale_y_continuous(labels = scales::label_dollar()) + 
    geom_point(alpha = 0.5) +
    geom_smooth(method = "lm", se = FALSE) +
    labs(
      subtitle = "Tuition vs. Full Time Faculty",
      x = "Percentage Full Time Faculty",
      y = "In-State Tuition",
      color = "Insitution Type"
    ) + theme_bw()


pctpellscp <- data %>%
  select(PCTPELL, TUITIONFEE_IN, CONTROL) %>% 
  drop_na() %>%
  ggplot(., aes(x = PCTPELL, y = TUITIONFEE_IN,  color = CONTROL)) +
    scale_x_continuous(labels = scales::label_percent()) + 
    scale_y_continuous(labels = scales::label_dollar()) + 
    geom_point(alpha = 0.5) +
    geom_smooth(method = "lm", se = FALSE) +
    labs(
      subtitle = "Tuition vs. Percent Pell Grant",
      x = "Pell Grant Rate",
      y = "In-State Tuition",
      color = "Insitution Type"
    ) + theme_bw() 

pctflscp <- data %>%
  select(PCTFLOAN, TUITIONFEE_IN, CONTROL) %>% 
  drop_na() %>%
  ggplot(., aes(x = PCTFLOAN, y = TUITIONFEE_IN,  color = CONTROL)) +
    scale_x_continuous(labels = scales::label_percent()) + 
    scale_y_continuous(labels = scales::label_dollar()) + 
    geom_point(alpha = 0.5) +
    geom_smooth(method = "lm", se = FALSE) +
    labs(
      subtitle = "Tuition vs. Percent Student Loan",
      x = "Student Loan Rate",
      y = "In-State Tuition",
      color = "Insitution Type"
    ) + theme_bw()

retft4scp <- data %>%
  select(RET_FT4, TUITIONFEE_IN, CONTROL) %>% 
  drop_na() %>%
  ggplot(., aes(x = RET_FT4, y = TUITIONFEE_IN,  color = CONTROL)) +
    scale_x_continuous(labels = scales::label_percent()) + 
    scale_y_continuous(labels = scales::label_dollar()) + 
    geom_point(alpha = 0.5) +
    geom_smooth(method = "lm", se = FALSE) +
    labs(
      subtitle = "Tuition vs. Rate First Time Undergrads",
      x = "First Time Undergrad Rate",
      y = "In-State Tuition",
      color = "Insitution Type"
    ) + theme_bw() 

gdmfcp <- data %>%
  select(GRAD_DEBT_MDN, TUITIONFEE_IN, CONTROL) %>% 
  drop_na() %>%
  ggplot(., aes(x = GRAD_DEBT_MDN, y = TUITIONFEE_IN,  color = CONTROL)) +
    scale_x_continuous(labels = scales::label_dollar()) + 
    scale_y_continuous(labels = scales::label_dollar()) + 
    geom_point(alpha = 0.5) +
    geom_smooth(method = "lm", se = FALSE) +
    labs(
      subtitle = "Tuition vs. Median Graduation Debt",
      x = "Median Graduation Debt",
      y = "In-State Tuition",
      color = "Insitution Type"
    ) + theme_bw()
admscp + ugdsscp + avgfacsalscp + pftfacscp + pctpellscp + pctflscp + retft4scp + gdmfcp + plot_layout(ncol = 2, guides = "collect") & theme(legend.position="bottom")
```

We also compared the `TUITIONFEE_IN` against some of the categorical variables in our data: `SINGLEGENDER`, `RELIGIOULS`, and `LOCALE`(@fig-cost-by-cat). It looks like there may be some significant difference in the distribution of `TUITIONFEE_IN` based on `SINGLE_GENDER` and `RELIGIOUS`. It's difficult to discern a major difference in `TUITIONFEE_IN` across the different levels of `LOCALE`.

```{r, echo=FALSE}
#| echo: false
#| label: fig-cost-by-cat
#| fig-cap: "Tuition versus Categorical Variables"
gendviol <- data %>%
  select(SINGLEGENDER, TUITIONFEE_IN, CONTROL) %>% 
  drop_na() %>%
  ggplot(., aes(x = SINGLEGENDER, y = TUITIONFEE_IN)) +
    geom_violin() +
    labs(
      subtitle = "Tuition vs. Single Gender Schools",
      x = "Single Gender",
      y = "In-State Tuition",
    ) + theme_bw()

relviol <- data %>%
  select(TUITIONFEE_IN, CONTROL, RELIGIOUS) %>% 
  drop_na() %>%
  ggplot(., aes(x = RELIGIOUS, y = TUITIONFEE_IN)) +
    geom_violin() +
    labs(
      subtitle = "Tuition vs. Religous Schools",
      x = "Single Gender",
      y = "In-State Tuition",
    ) + theme_bw()

locviol <- data %>%
  select(LOCALE, TUITIONFEE_IN, CONTROL) %>% 
  drop_na() %>%
  ggplot(., aes(x = LOCALE, y = TUITIONFEE_IN)) +
    geom_violin() +
    labs(
      subtitle = "Tuition vs. School Locale",
      x = "Locale",
      y = "In-State Tuition",
    ) + theme_bw()

gendviol + relviol + locviol + plot_layout(ncol = 2)
```

## Model Building

### Initial Model

We know we want to include interaction terms with CONTROL and other predictors. Because of that we know that we will not be able to rely on the automated search procedures we've used previously. So the initial model will start off as a first order model  including predictors identified from the charts shown previously. These predictors are:

-   `ADM_RATE` - based on scatter plot
-   `AVGFACSAL` - based on scatter plot
-   `PCTPELL` - based on scatter plot
-   `RET_FT4` - based on scatter plot
-   `CONTROL` - based on scatter plots and hypothesis that interaction terms will be important

`PCFTLOAN` was not included because it looks to be correlated with `PCTPELL`.

```{r, echo=FALSE}
initmodel <- lm(TUITIONFEE_IN~ADM_RATE+AVGFACSAL+PCTPELL+RET_FT4+CONTROL, data = data)
summary(initmodel)
```

### Improved Model

After creating the initial first order model based on observations from the plots displayed earlier we will compare it to a model that contains the remaining predictors we obtained data on to ensure we weren't missing any important ones.

```{r, echo=FALSE}
largemodel <- lm(TUITIONFEE_IN~ADM_RATE+AVGFACSAL+PCTPELL+RET_FT4+CONTROL+UGDS+INEXPFTE+
                   PFTFAC+PCTFLOAN+GRAD_DEBT_MDN+RELIGIOUS+SINGLEGENDER+LOCALE, data = data)
summary(largemodel)
```

The adjusted $R^2$ metric as well as the results of the t-tests for some of coefficients for the additional predictors hint that they may be significant for our model. To confirm these additional predictors are useful we will run a partial F-test where:
$$H_0: \beta_2 = 0$$
$$H_a: \beta_2 \neq 0$$
where $\beta_2$ is the collection of coefficients added in the expanded model.

```{r, echo=FALSE}
anova(initmodel,largemodel)
```

The p-value form the partial F-test is less than 0.05 indicating that the data supports rejecting the null hypothesis which indicates at least some of our additional predictors add value to the model. However some of the t-tests indicate that we may be able to remove some of the predictors namely:

* `PFTFAC`
* `SINGLEGENDER`
* `LOCALE`

Again, to confirm we can conduct an additional F-test. The hypothesis test remains the same as well except with $B_2$ representing the coefficients `PFTFAC`, `SINGLEGENDER`, and `LOCALE`.
$$H_0: \beta_2 = 0$$
$$H_a: \beta_2 \neq 0$$

```{r, echo=FALSE}
finfirstmodel <- lm(TUITIONFEE_IN~ADM_RATE+AVGFACSAL+PCTPELL+RET_FT4+CONTROL+UGDS+INEXPFTE+
                   PCTFLOAN+GRAD_DEBT_MDN+RELIGIOUS, data = data)
summary(finfirstmodel)
```

```{r, echo=FALSE}
anova(finfirstmodel,largemodel)
```

For this partial F-test the p-value is greater than 0.05 indicating they the data does not support rejecting the null hypothesis. Which indicates that the predictors `PFTFAC`, `SINGLEGENDER`, and `LOCALE` were not significant.

After creating a first order model we will attempt to improve it using interaction terms with CONTROL. The interaction terms were selected based off of the scatter plots shown previously.

```{r, echo=FALSE}
intmodel1 <- lm(TUITIONFEE_IN~ADM_RATE+AVGFACSAL+PCTPELL+RET_FT4+CONTROL
                +UGDS+INEXPFTE+PCTFLOAN+GRAD_DEBT_MDN+RELIGIOUS+
                  ADM_RATE:CONTROL+AVGFACSAL:CONTROL+PCTPELL:CONTROL+
                  RET_FT4:CONTROL, data = data)
summary(intmodel1)
```

Once again we will conduct a partial F test to determine if the additional terms in the model are significant.
$$H_0: \beta_2 = 0$$
$$H_a: \beta_2 \neq 0$$
where $\beta_2$ represents the coefficients for the interaction terms.

```{r, echo=FALSE}
anova(finfirstmodel,intmodel1)
```

The p-value for the partial F test is less than 0.05 so the data supports rejecting the null hypothesis. This indicates that some of the interaction terms are significant to the model.

The t-test indicates that the interaction term between `RET_FT4` and the indicator variable for `CONTROLPrivate` can be dropped from the model given that the rest are present. This gives us the model below.

```{r, echo=FALSE}
intmodel2 <- lm(TUITIONFEE_IN~ADM_RATE+AVGFACSAL+PCTPELL+RET_FT4+CONTROL+UGDS+INEXPFTE+
                   PCTFLOAN+GRAD_DEBT_MDN+RELIGIOUS+
                  ADM_RATE:CONTROL+AVGFACSAL:CONTROL+PCTPELL:CONTROL, data = data)
summary(intmodel2)
```

Before we determine if this is the recommended model we should assess the regression assumptions.

```{r, echo=FALSE}
tuithat <- intmodel2$fitted.values
res <- intmodel2$residuals
data <- data.frame(data,tuithat,res)
```

```{r, echo=FALSE}
#| fig-cap: "Residuals versus Fitted for intermediate model 2"
#| label: fig-residual-model-2
ggplot(data, aes(x=tuithat,y=res))+
  geom_point()+
  geom_hline(yintercept=0, color="red")+
  labs(x="Fitted y", y="Residuals", title="Residual Plot") +
  theme_bw()
```

The residuals (@fig-residual-model-2) look to have a mean of 0 for most of the fitted values of y but at the higher fitted values of y there looks to be an area where the mean is less then 0. In addition the variance of the residuals looks to increase as the fitted values increase. So we do not meet the first 2 regression assumptions. To attempt to correct for this we will transform the response.

```{r, echo=FALSE}
#| fig-cap: "Box-Cox plot for intermediate model 2"
#| label: fig-boxcox
boxcox(intmodel2, lambda = seq(-1,1,1/10))
```

The Box-Cox plot (@fig-boxcox) indicates a $\lambda$ somewhat near 0. Because of that and to aid in interpretation we will use the log transformation.

```{r, echo=FALSE}
logtuit <- log(data$TUITIONFEE_IN)
data <- data.frame(data,logtuit)
```

```{r, echo=FALSE}
logmodel <- lm(logtuit~ADM_RATE+AVGFACSAL+PCTPELL+RET_FT4+CONTROL+UGDS+INEXPFTE+
                   PCTFLOAN+GRAD_DEBT_MDN+RELIGIOUS+
                  ADM_RATE:CONTROL+AVGFACSAL:CONTROL+PCTPELL:CONTROL, data = data)
summary(logmodel)
```

```{r, echo=FALSE}
logtuithat <- logmodel$fitted.values
logres <- logmodel$residuals
data <- data.frame(data,logtuithat,logres)
```

```{r, echo=FALSE}
#| fig-cap: "Residuals vs Fitted for log transformed model"
#| label: fig-residual-log 
ggplot(data, aes(x=logtuithat,y=logres))+
  geom_point()+
  geom_hline(yintercept=0, color="red")+
  labs(x="Fitted y", y="Residuals", title="Residual Plot") +
  theme_bw()
```

With the log transformation on `TUITIONFEE_IN` the variance of the residuals does look to be fairly constant for the fitted values (@fig-residual-log). There is a chance at high fitted values (above 11) that there is an area where the mean of the residuals is not 0, but mostly it looks fine.

```{r, echo=FALSE}
#| fig-cap: "Autocorrelation plot of residuals for log transformed model"
#| label: fig-acf-log 
acf(logres, main="ACF Plot of Residuals with Log Tuition")
```

It appears that the residuals do have some correlation with each other as at a few different lags (@fig-acf-log). It looks like this could possibly be because the data has been sorted by the school names. Many schools are named "University of <State>" and are the flagship school for their state; these schools are likely correlated. It's reasonable to make the assumption that the autocorrelation of the residuals is likely a result of how the data has been sorted.

```{r, echo=FALSE}
#| fig-cap: "Q-Q Plot for the log transformed model"
#| label: fig-qq-log
qqnorm(logres)
qqline(logres, col="red")
```

For the most part, the residuals look to follow the normal distribution (@fig-qq-log), at the lower end there is some deviation, but not enough to be a major concern.

From the 3 charts above we can conclude that the model comes close enough to meeting the regression assumptions.

We used Cook's distance and DFFITS to assess influential observations. Cooks distance did not indicate that there were any influential observations. However DFFITS indicated that there were 84 influential observations. Below we built the linear model after removing the influential data points.

```{r, echo=FALSE}
n <- nrow(data)
p <- 14
COOKS <- cooks.distance(logmodel)
datacooks <- data[COOKS>qf(0.5,p,n-p),]

DFFITS<-dffits(logmodel)
data$DFFITS <- DFFITS

data2 <- data %>% 
  filter(abs(DFFITS) <= 2*sqrt(p/n))

dffitsmodel <- lm(logtuit~ADM_RATE+AVGFACSAL+PCTPELL+RET_FT4+CONTROL+UGDS+INEXPFTE+
                   PCTFLOAN+GRAD_DEBT_MDN+RELIGIOUS+
                  ADM_RATE:CONTROL+AVGFACSAL:CONTROL+PCTPELL:CONTROL, data = data2)

influential_points <- data %>%
  mutate(adm_q = ecdf(ADM_RATE)(ADM_RATE), ret_q = ecdf(RET_FT4)(RET_FT4)) %>%
  filter(abs(DFFITS) > 2*sqrt(p/n)) %>%
  select(INSTNM, ADM_RATE, adm_q, ret_q, RET_FT4)

summary(dffitsmodel)
```

The most notable difference is that the t-tests for `RET_FT4` and `ADM_RATE` indicates that both may not be a important to the model given the presence of the other predictors without the influential data points. The estimated coefficients for the other predictors also changed but there were not changes in the signs. If we examine these influential points, we notice that most of these schools are on the extremes, including:

* Ivy League schools (Harvard, Stanford, Princeton) with a very low admission rate (around 5%, less than 1st percentile)
* Schools with near 100% admission rates (Moody Bible Institute, Prescott College, Middle Georgia State University, Unity College, Criswell College)
* Two schools with a 0% retention rate (San Francisco Art Institute, Cleveland University-Kansas City)
* Schools with a >99% retention rate. Among these schools, two are moderately competitive (University of Virginia-Main Campus, University of Florida), two are highly competitive (Massachusetts Institute of Technology, University of Chicago), and one is not competitive (Platt College-Aurora).

In summary, `ADM_RATE` and `RET_FT4` are correlated at the extremes and removing these points removes the observed association between these predictors and our outcome. This indicates that our estimates of the coefficients for `ADM_RATE` and `RET_FT4` are not robust in terms of influential data points.

### Recommended Model

The recommended model is below. It includes a log transform on the response variable to account for non-constant variation that was seen in the previous model. The t-test for `ADM_RATE`indicates that it could be removed from the model leaving in all other regressors but it needs to be included because there is an interaction term between `ADM_RATE` and `CONTROL` that is included in the model.

```{r, echo=FALSE}
summary(logmodel)
```

Based on both the Adjusted $R^2$ and AIC metrics (@tbl-model-metrics) the recommended model is considered the best model out of the ones we have tested. The agreement between the two measure increases our confidence in the selection of the recommended model.

|                                      | Adjusted $R^2$ | AIC      |
|--------------------------------------|----------------|----------|
|  Initial First Order Model           | 0.8013         | 29299.02 |
| First Order Model with All Terms     |  0.8324        | 29067.13 |
| Final First Order Model              | 0.8324         | 29062.05 |
| Model with Interaction Terms         | 0.8586         | 28822.47 |
| Recommended Model                    | 0.8587         | 28822.47 |
| Recommended Model with Log Transform | 0.8809         | 95.94833 |

: Fit metrics for selected models {#tbl-model-metrics}

```{r,include=FALSE}
print(AIC(initmodel))
print(AIC(largemodel))
print(AIC(finfirstmodel))
print(AIC(intmodel1))
print(AIC(intmodel2))
print(AIC(logmodel))
```

## Conclusions

Based on our linear model we can conclude that the following predictors have a relationship with the cost of in-state tuition at U.S. universities: `AVGFACSAL`, `PCTPELL`, `RET_FT4`, `CONTROL`, `UGDS`, `PCTFLOAN`, `GRAD_DEBT_MDN`, and `RELIGIOUS`. We can also conclude that the following predictors have a different relationship to the cost of in-state tuition when `CONTROL` is public vs. private: `ADM_RATE`, `AVGFAC_SAL`, and `PCTPELL`. It should be noted that the estimated coefficients for `RET_FT4` and `ADM_RATE` are not robust to influential data points so we will delay drawing conclusions from those predictors.

We can make the following conclusions from our analysis:

-   **Tuition at private schools is likely to be higher than public schools.** All other variables held constant a private schools estimated in-state tuition is 6.9797 ($e^{1.943}$) times that of a public school.

-   **High instructional expenditures are associated with higher tuition.** All other variables held constant a 1,000 dollar increase in instructional expenditures per full time student multiples the estimated in-state tuition by 1.0045 ($e^{4.468 \times 10^{-6} \times 1000}$).

-   **Schools with higher in-state tuition have students that take out more in loans.** All other variables held constant a 1,000 dollar increase in median student debt at graduation multiples the estimated in-state tuition by 1.0179 ($e^{1.772 \times 10^{-5} \times 1000}$).

-   **Schools with a high percentage of Pell grant recipients have lower tuition.** For public schools all other variables held constant a 1% increase in Pell grant recipients multiplies the estimated tuition by 0.9939 ($e^{-6.058 \times 10^{-1} \times 0.01}$). For private schools all other variables held constant a 1% increase in Pell grant recipients multiples the estimated tuition by 0.9849 ($e^{-6.058 \times 10^{-1} - 9.167 \times 10^{-1} \times 0.01}$).

In summary we found that higher values for the following predictors are associated with higher estimated in-state tuition given other predictors are held constant: average faculty salary, instructional expenditure per student, proportion of students who received a federal student loan, and median student debt at graduation. We found that higher values for the following predictors proportion of students receiving Pell grants and the number of undergraduate degree seeking students are associated with lower estimated in-state tuition. Private schools are associated with in-state tuition when compared to public schools given other predictors are held constant. Religious schools are associated with lower in-state tuition compare to secular schools all other predictors held constant. We found that the relationship between estimated in-state tuition differs for average faculty salary and proportion of Pell grant recipients when the school is private versus when they are public.
