---
title: "Group 5 Proposal"
author:
  - "Andrew Barros"
  - "Audrika Indu"
  - "Sivaranjani Kandasami"
  - "Alexander Kendrick"
format: pdf
---

```{r}
#| message: false
#| code-fold: true
library(tidyverse)
data <- read_csv(
  "../data/analysis_data.csv",
  col_types = cols(
    UNITID = col_double(),
    CDR3 = col_double(),
    INSTNM = col_character(),
    HBCU = col_double(),
    MENONLY = col_double(),
    WOMENONLY = col_double(),
    RELAFFIL = col_double(),
    DISTANCEONLY = col_double(),
    CONTROL = col_character(),
    LOCALE = col_character(),
    ADM_RATE = col_double(),
    SATVRMID = col_double(),
    SATMTMID = col_double(),
    SATWRMID = col_double(),
    UGDS = col_double(),
    TUITIONFEE_IN = col_double(),
    TUITIONFEE_OUT = col_double(),
    INEXPFTE = col_double(),
    AVGFACSAL = col_double(),
    PFTFAC = col_double(),
    PCTPELL = col_double(),
    PCTFLOAN = col_double(),
    RET_FT4 = col_double(),
    GRAD_DEBT_MDN = col_double(),
    HIGH_CDR = col_logical(),
    SINGLEGENDER = col_double(),
    RELIGIOUS = col_logical()
  )
)
```

# Data Set

Our data comes from the [US Department of Education College Scorecard](https://collegescorecard.ed.gov/), specifically the ["Most Recent Institution-Level Data" from September 14, 2022](https://ed-public-download.app.cloud.gov/downloads/Most-Recent-Cohorts-Institution_09012022.zip). The raw dataset contains 2,989 features on 6,681 institutions but we've chosen to focus on 1,990 insitutions that are:

-   Primarily four-year degree granting insitutions (for undergraduates)
-   The main campus only (if an overall insitutition has mulitple campuses)

[This dataset is well documented](https://collegescorecard.ed.gov/data/documentation/) and [we have built a small pipeline for ingesting the raw DOE data and creating a dataset for our analysis](https://github.com/ajb5d/CollegeScorecard/blob/main/get_data.R). Our proposed dataset meets the requirements of 2.1.(b), specifically:

-   This data is not from the textbook
-   We have not previously used this data in class
-   None of our group members have worked with this dataset previously
-   It is not time-related data
-   It is real-world (not simulated) data

# Questions of Interest

## Question of Interest 1: What influences tuition?

In this aim, we will examine the factors influencing the total in-state tuition and fees for main campus, four year, degree granting insitutions.

Our primary response variable is `TUITIONFEE_IN` -- the total tuition and fees for in-state students in US dollars. We hypothesize that the most important predictors will be:

-   `CONTROL` -- the type of insitution (public, private/non-profit, private/for-profit)
-   `LOCALE` -- the locale of the campus (categorical, nine levels from City: Large to Rural: Remote)
-   `DISTANCEONLY` -- an indicator if the insitution is online-only
-   `RELIGIOUS` -- an indicator if the institution is relgiously affiliated
-   `SINGLEGENDER` -- an indicator if the insitution is single-gender
-   `HBCU` -- an indicator if the instution is a historically Black college or university (HBCU)
-   `ADM_RATE` -- the admission rate
-   `AVGFACSAL` -- the average faculty salary
-   `PFTFAC` -- the percentage of full-time faculty
-   `UGDS` -- the number of undergraduate degree and certificate seeking students
-   `INEXPFTE` -- the instructional expenditures per full time student
-   `PCTPELL` -- the percentage of students recieving a Pell grant
-   `RET_FT4` -- First-time, full-time student retention rate

**FIXME**: For each question, write a short paragraph to explain why the question is worth exploring. The more interesting, the better.


## Question of Interest 2: What is associated with having a high student loan default rate?

In this aim, we will examine the factors associated with having a three-year cohort federal student loan default rate higher than 10% for main campus, four year, degree granting insitutions.

Our primary response variable is `HIGH_CDR` -- a indicator variable we created when the three year cohort default rate (CDR3) is greater than 10%. We hypothesize that the most important predictors will be:

-   `CONTROL` -- the type of insitution (public, private/non-profit, private/for-profit)
-   `LOCALE` -- the locale of the campus (categorical, nine levels from City: Large to Rural: Remote)
-   `DISTANCEONLY` -- an indicator if the insitution is online-only
-   `RELIGIOUS` -- an indicator if the institution is relgiously affiliated
-   `SINGLEGENDER` -- an indicator if the insitution is single-gender
-   `HBCU` -- an indicator if the instution is a historically Black college or university (HBCU)
-   `TUITIONFEE_IN`
-   `ADM_RATE` -- the admission rate
-   `UGDS` -- the number of undergraduate degree and certificate seeking students
-   `INEXPFTE` -- the instructional expenditures per full time student
-   `PCTPELL` -- the percentage of students recieving a Pell grant
-   `RET_FT4` -- First-time, full-time student retention rate
-   `GRAD_DEBT_MDN` -- the median debt at graduation
-   `PCTFLOAN` -- the percentage of students graduating with federal loans

**FIXME**: For each question, write a short paragraph to explain why the question is worth exploring. The more interesting, the better.

# Visualizations
## Question of Interest 1
```{r, echo=FALSE}
ggplot(data, aes(x = ADM_RATE, y = TUITIONFEE_IN, color = CONTROL))+
  geom_point(alpha = 0.5)+
  labs(title = "Admission Rate vs. Tuition and Control", x = "Admission Rate", y = "Tuition", color = "Control")
```

```{r, echo=FALSE}
ggplot(data, aes( x = AVGFACSAL, y = TUITIONFEE_IN,  color = CONTROL))+
  geom_point(alpha = 0.5)+
  labs(title = "Admission Rate vs. Faculty Salary and Control", x = "Faculty Salary", y = "Tuition", color = "Control")
```

```{r, echo=FALSE}
ggplot(data, aes( x = SATVRMID, y = TUITIONFEE_IN,  color = CONTROL))+
  geom_point(alpha = 0.5)+
  labs(title = "Admission Rate vs. Faculty Salary and Control", x = "SAT Read", y = "Tuition", color = "Control")
```
```{r, echo=FALSE}
ggplot(data, aes( x = SATMTMID, y = TUITIONFEE_IN,  color = CONTROL))+
  geom_point(alpha = 0.5)+
  labs(x = "SAT Read", y = "Tuition", color = "Control")
```


## Question of Interest 2
```{r}
ggplot(data = data, aes(x = HIGH_CDR))+
  geom_bar()+
  labs(title = "High Student Loand Default",x = "High Student Loan Default", y = "Count", fill = "Control")
```

```{r}
ggplot(data = data, aes(x = HIGH_CDR, fill = factor(HBCU)))+
  geom_bar()+
  labs(title = "High Student Loand Defaults for HBCUs", x = "High Student Loan Default", y = "Count", fill = "HBCU")
```

```{r}
ggplot(data = data, aes(x = HIGH_CDR, y = GRAD_DEBT_MDN, fill = CONTROL))+
  geom_boxplot()+
  labs(title = "Student Debt at Grad vs. High Student Loan Default and Control",x = "High Student Loan Default", y = "Graduation Debt", fill = "Control")
```

```{r}
ggplot(data = data, aes(x = HIGH_CDR, y = SATMTMID, fill = CONTROL))+
  geom_boxplot()+
  labs(title = "Student Debt at Grad vs. Undergrads and Control",x = "High Student Loan Default", y = "Undergrads", fill = "Control")
```



