---
title: "Group 5 Proposal"
author:
  - "Andrew Barros"
  - "Audrika Indu"
  - "Sivaranjani Kandasami"
  - "Alexander Kendrick"
format: pdf
---

```{r}
#| message: false
#| echo: false
library(tidyverse)

data <- read_csv(
  "../data/analysis_data.csv",
  col_types = cols(
    UNITID = col_double(),
    CDR3 = col_double(),
    INSTNM = col_character(),
    HBCU = col_double(),
    MENONLY = col_double(),
    WOMENONLY = col_double(),
    RELAFFIL = col_double(),
    DISTANCEONLY = col_double(),
    CONTROL = col_character(),
    LOCALE = col_character(),
    ADM_RATE = col_double(),
    SATVRMID = col_double(),
    SATMTMID = col_double(),
    SATWRMID = col_double(),
    UGDS = col_double(),
    TUITIONFEE_IN = col_double(),
    TUITIONFEE_OUT = col_double(),
    INEXPFTE = col_double(),
    AVGFACSAL = col_double(),
    PFTFAC = col_double(),
    PCTPELL = col_double(),
    PCTFLOAN = col_double(),
    RET_FT4 = col_double(),
    GRAD_DEBT_MDN = col_double(),
    HIGH_CDR = col_logical(),
    SINGLEGENDER = col_double(),
    RELIGIOUS = col_logical()
  )
)
```

# Data Set

Our data comes from the [US Department of Education College Scorecard](https://collegescorecard.ed.gov/), specifically the ["Most Recent Institution-Level Data" from September 14, 2022](https://ed-public-download.app.cloud.gov/downloads/Most-Recent-Cohorts-Institution_09012022.zip). The raw dataset contains 2,989 features on 6,681 institutions but we've chosen to focus on 1,990 institutions that are:

-   Primarily four-year degree granting institutions (for undergraduates)
-   The main campus only (if an overall insitutition has mulitple campuses)

[This dataset is well documented](https://collegescorecard.ed.gov/data/documentation/) and [we have built a small pipeline for ingesting the raw DOE data and creating a dataset for our analysis](https://github.com/ajb5d/CollegeScorecard/blob/main/get_data.R). Our proposed dataset meets the requirements of 2.1.(b), specifically:

-   This data is not from the textbook
-   We have not previously used this data in class
-   None of our group members have worked with this dataset previously
-   It is not time-related data
-   It is real-world (not simulated) data

# Questions of Interest

## Question of Interest 1: What influences tuition?

In this aim, we will examine the factors influencing the total in-state tuition and fees for main campus, four year, degree granting institutions.

Our primary response variable is `TUITIONFEE_IN` -- the total tuition and fees for in-state students in US dollars. We hypothesize that the most important predictors will be:

-   `CONTROL` -- the type of institution (public, private/non-profit, private/for-profit)
-   `LOCALE` -- the locale of the campus (categorical, nine levels from City: Large to Rural: Remote)
-   `DISTANCEONLY` -- an indicator if the institution is online-only
-   `RELIGIOUS` -- an indicator if the institution is relgiously affiliated
-   `SINGLEGENDER` -- an indicator if the institution is single-gender
-   `HBCU` -- an indicator if the instution is a historically Black college or university (HBCU)
-   `ADM_RATE` -- the admission rate
-   `AVGFACSAL` -- the average faculty salary
-   `PFTFAC` -- the percentage of full-time faculty
-   `UGDS` -- the number of undergraduate degree and certificate seeking students
-   `INEXPFTE` -- the instructional expenditures per full time student
-   `PCTPELL` -- the percentage of students recieving a Pell grant
-   `RET_FT4` -- First-time, full-time student retention rate

The cost of attending college is one of the main concerns for high school students planning for undergraduate studies. Tuition and fees have significantly increased over the past decade which resulted in increased student debts and an increase in student loans: 

* Total $1.75 trillion in student loan debt (including federal and private loans)
* On average, each borrower owes $28,950 
* About 92% of all student debt are federal student loans
* More than half (55%) of students from public four-year institutions had student loans

With the above facts, it is important that we understand the factors that influence high tuition fees. These factors that impact the fees needs to be addressed to encourage the current and future generations can have a debt-free college education of their choice. 

References:

* https://studentloanhero.com/featured/undergrad-worries-survey/
* https://educationdata.org/student-loan-debt-crisis

## Question of Interest 2: What is associated with having a high student loan default rate?

In this aim, we will examine the factors associated with having a three-year cohort federal student loan default rate higher than 10% for main campus, four year, degree granting institutions.

Our primary response variable is `HIGH_CDR` -- a indicator variable we created when the three year cohort default rate (CDR3) is greater than 10%. We hypothesize that the most important predictors will be:

-   `CONTROL` -- the type of institution (public, private/non-profit, private/for-profit)
-   `LOCALE` -- the locale of the campus (categorical, nine levels from City: Large to Rural: Remote)
-   `DISTANCEONLY` -- an indicator if the institution is online-only
-   `RELIGIOUS` -- an indicator if the institution is religiously affiliated
-   `SINGLEGENDER` -- an indicator if the institution is single-gender
-   `HBCU` -- an indicator if the institution is a historically Black college or university (HBCU)
-   `TUITIONFEE_IN`
-   `ADM_RATE` -- the admission rate
-   `UGDS` -- the number of undergraduate degree and certificate seeking students
-   `INEXPFTE` -- the instructional expenditures per full time student
-   `PCTPELL` -- the percentage of students receiving a Pell grant
-   `RET_FT4` -- First-time, full-time student retention rate
-   `GRAD_DEBT_MDN` -- the median debt at graduation
-   `PCTFLOAN` -- the percentage of students graduating with federal loans

As tuition costs and student loans increase a key factor in a student deciding if they will attend college is whether they will see a return on their investment. High student loan default rate can indicate that students aren't receiving an adequate career return on their investment after attending college. It is important to determine what factors influence loan repayment so that colleges and students can analyze what will allow them to get the most out of their education.

# Visualizations
## Question of Interest 1: What Influences Tuition Cost?
```{r, echo=FALSE}
data %>%
  select(ADM_RATE, TUITIONFEE_IN, CONTROL) %>% 
  drop_na() %>% 
  ggplot(aes(x = ADM_RATE, y = TUITIONFEE_IN, color = CONTROL)) +
    scale_x_continuous(labels = scales::label_percent()) + 
    scale_y_continuous(labels = scales::label_dollar()) + 
    geom_point(alpha = 0.5) +
    labs(
      title = "Admission Rate vs. Tuition and institution type",
      x = "Admission Rate",
      y = "In-State Tuition",
      color = "institution Type"
    ) + theme_bw()
```

```{r, echo=FALSE}
data %>%
  select(AVGFACSAL, TUITIONFEE_IN, CONTROL) %>% 
  drop_na() %>%
  ggplot(., aes(x = AVGFACSAL, y = TUITIONFEE_IN,  color = CONTROL)) +
    scale_x_continuous(labels = scales::label_dollar()) + 
    scale_y_continuous(labels = scales::label_dollar()) + 
    geom_point(alpha = 0.5) +
    labs(
      title = "Tuition vs. Average Faculty Salary by institution type",
      x = "Average Faculty Salary (USD per month)",
      y = "In-State Tuition",
      color = "institution Type"
    ) + theme_bw()
```

```{r, echo=FALSE}

data %>%
  select(SATVRMID:SATWRMID, TUITIONFEE_IN, CONTROL) %>%
  pivot_longer(SATVRMID:SATWRMID) %>%
  mutate(name = case_when(
    name == "SATVRMID" ~ "Verbal",
    name == "SATMTMID" ~ "Math",
    name == "SATWRMID" ~ "Writing",
  )) %>%
  filter(!is.na(value), !is.na(TUITIONFEE_IN)) %>% 
  ggplot(aes(x = value, y = TUITIONFEE_IN, color = CONTROL)) + 
    geom_point(alpha = 0.5) +
    scale_y_continuous(labels = scales::label_dollar()) +
    labs(
      title = "Median SAT Score vs In-State Tuition by institution type",
      y = "In-State Tutition",
      y = "Median SAT Score by Section", 
      color = "Institution Type"
    ) + 
    facet_wrap(~name) +
    theme_bw() + 
    theme(legend.position='bottom')
```


## Question of Interest 2: What is associated with having a high student loan default rate?

```{r}
#| echo: false

data %>%
  filter(!is.na(HIGH_CDR)) %>%
  ggplot(aes(x = HIGH_CDR)) +
    geom_bar() +
    labs(title = "High Student Loan Default",
         x = "High Student Loan Default",
         y = "Count",
         fill = "Control") +
    theme_bw()
```

```{r}
#| echo: false
data %>%
  filter(!is.na(TUITIONFEE_IN), !is.na(UGDS), !is.na(HIGH_CDR)) %>%
  ggplot() +
    geom_violin(aes(HIGH_CDR, UGDS, fill = HIGH_CDR)) +
    scale_y_continuous(labels = scales::label_comma(), trans = "log10") +
    labs(title = "Undergrad Population Size vs. High Student Loan Default Rate and Control",
         x = "High Student Loan Default",
         y = "Undergraduate Population Size",
         title = "UG Pop vs Institution Type by High Student Loan Default",
         color = NULL) + 
    facet_wrap(~CONTROL) +
    guides(fill = 'none') + 
    theme_bw()
```


```{r}
#| echo: false
data %>%
  filter(!is.na(HIGH_CDR), !is.na(GRAD_DEBT_MDN)) %>%
  ggplot(aes(x = HIGH_CDR, y = GRAD_DEBT_MDN, fill = CONTROL)) +
  geom_violin() +
  scale_y_continuous(labels = scales::label_dollar()) +
  labs(title = "Student Debt vs. High Student Loan Default and institution type",
       x = "High Student Loan Default",
       y = "Median Graduation Debt",
       fill = "Control") +
  theme_bw()
```


